---
title: "Task5"
author: "Gina Miku Oba"
date: "2023-03-06"
output:
  html_document:
    df_print: paged
---

# Task5 - Using R example datasets

## 1. CO2 dataset
First, load CO2 dataset.
```{r loading}
data(CO2)
# also loading library for later use
library(tidyverse)
```

This data contains 84 observations (row) and 5 variables (column). 

## 2. Contents of CO2 dataset
```{r}
help(CO2)
```

The CO2 dataset is from an experiment on the cold tolerance of the grass species _Echinochloa crus-galli_ from Quebec and Mississippi. This data contains 84 observations (row) and 5 variables (column). There are 3 biological replicates for each 4 condition (Quebec-nonchilled, Quebec-chilled, Mississippi-nonchilled, Mississippi-chilled) and 7 different ambient carbon dioxide concentration for each sample (3 x 4 x 7 = 84). 5 variables include the unique identifier for each plant (*Plant*), the origin of the plant (*Type*),  treatment (*Treatment*, which is `nonchilled` or `chilled`), the ambient carbon dioxide concentration (*conc*), the carbon dioxide uptake rates (*uptake*) of the plant.

## 3. Average and median CO2 uptake of the plants from Quebec and Misissippi.
Functions to compute Average and median in R is as below.

- average : `mean(x)`
- median : `median(x)`

```{r}
# average(mean) and median CO2 uptake of the all plants
CO2 %>%
  # summarize results in table form
    summarize( 
      Average = mean(uptake, na.rm=TRUE),
      Median = median(uptake, na.rm=TRUE))
```

Compute average(mean) and median CO2 uptake of the plants from Quebec and Mississippi.
```{r}
# average(mean) and median CO2 uptake of the plants from Quebec and Mississippi.
CO2 %>%
    group_by(Type) %>% # grouping data of same origin
    # summarize results in table form
    summarize(
      Average = mean(uptake, na.rm=TRUE),
      Median = median(uptake, na.rm=TRUE))
```
From the data, we can speculate that the plant from Quebec will have more CO2 uptake than the plant from Mississippi.

## 4. [Optional] How many genes are expressed in each sample in "airway" dataset? How many genes are not expressed in any sample?

To load airway dataset from Bioconductor, command is as below.

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("airway")

library( "airway" )
data("airway")
```

```{r}
help(airway)
airway
```
`airway` data is the RNA-seq read count data of airway smooth muscle cell lines, contains 64102 features of 8 samples. Matrix of read count (which row is gene, and column is sample) is able to extruct by `assay(airway)`.

If read count is over 1 that gene is considered as expressed gene.
```{r}
# count expressed gene (read count >=1) for each sample
as.data.frame(assay(airway)) %>% # change data format. Otherwise, it will be `Error in UseMethod("tbl_vars")`
  summarise_all(~sum(. >= 1, na.rm = TRUE)) # count if value is >=1.

```
Average of expressed gene across samples is as below.
```{r}
# count expressed gene (read count >=1) for each sample
as.data.frame(assay(airway)) %>% # change data format. Otherwise, it will be `Error in UseMethod("tbl_vars")`
  summarise_all(~sum(. >= 1, na.rm = TRUE)) %>% # count if value is >=1.
  # Mean of expressed gene across samples.
  rowMeans()
```

The gene which not expressed in any sample is as below.
```{r}
df = as.data.frame(assay(airway))

# rowSums() function to calculate the sum of each row
# filter the data frame with the row which sum of the row is 0.
all_zero_rows <- df[rowSums(df) == 0, ]
print(dim(all_zero_rows))

```

Across sample average 24767 genes, minimun 23124 to maximum 25998 genes are expressed (read count >=1).
And genes which is not expressed in any sample is 30633 genes.



