---
title: "Task6"
author: "Gina Miku Oba"
date: "2023-03-07"
output: html_notebook
---
# Task6 - R Function

## 1. A function that calculates the ratio of the mean and the median of a given vector.
Input of the function is given vector and output of the function is the ratio of the mean and the median.
$$
f(x) = \frac{mean(x)}{median(x)} \\
\text{where, }x = \{x_1, x_2, ..., x_n\}
$$

```{r}
ratio_mean_median <- function(x){
  return(mean(x) / median(x))
  }
```

Try test on random vector:
```{r}
# Generate a random vector of 10 components, each with a value between 0 and 1.
vector <- runif(10, 0, 1) 
print(vector)

# print mean and median
print(paste("mean:", mean(vector)))
print(paste("median:", median(vector)))

# print ratio of mean and median 
result <- ratio_mean_median(vector)
print(result)

# print if function works fine
if (result == mean(vector)/median(vector)){
  print("Function works well")
}else{
  print("Function not works")
}
```

## 2. A function that ignore the lowest and the hightest value from given vector and calculate the mean.
The input to the function is a vector. The output of the function is mean. The lowest value in the function is searched for with min() and the highest value with max().

```{r}
trim_mean <- function(x){
  min <- min(x)
  max <- max(x)
  # remove min and max from given vector using subset().
  trimmed <- subset(x, x != min & x != max)
  return(mean(trimmed))
}
```

Try test:
```{r}
# Generate a random vector of 10 components, each with a value between 0 and 1.
vector <- runif(10, 0, 1) 
print(vector)
print(paste("median:", median(vector)))

# print mean and trim_mean
print(paste("mean:", mean(vector)))
print(paste("trim_mean:", trim_mean(vector)))
```
The advantage of this trimmed mean is that it is robust to outliers or the heavy tail data distribution. 
Try vector with extreme value:
```{r}
# Generate a random vector of 10 components, each with a value between 0 and 1.
vector <- runif(9, 0, 1) 
# add extreme value 10
vector <- append(vector, 10)
print(vector)
print(paste("median:", median(vector)))

# print mean and trim_mean
print(paste("mean:", mean(vector)))
print(paste("trim_mean:", trim_mean(vector)))
```
By trimming the highest and lowest values, the trimmed mean was closer to the median.

## 3. Explanation of piping usase
Write a short (max. 300 characters, no spaces) explanation of why, how, and when not to use pipes.

Pipes chain functions together and improve code readability. However, longer chains (> 10) can complicate code and should be avoided. Pipes overwrite intermediate objects, making them unsuitable for functions with multiple inputs/outputs. Additionally, pipes have limitations for functions with complex dependencies since they operate linearly.

## 4. Explanation of apply-family function
Write a short explanation (max. 300 characters, no spaces) of why they could be useful in your work.

The Apply family of functions is useful for applying a function to entire dataset. Loops can create a deep hierarchy, making code difficult to read. Apply-functions can simplify code by reducing loops. They are crucial when analyzing large dataset such as gene expression matrices, streamlining the process and ensuring consistent results.